Tutorial: SQL-on-Hadoop Systems

### 2015-04-02 paper.tex
More refs

It is important to note that there are two important distinct characteristics of this emerging eco-system that is different than traditional relational warehouses. First, in the world of Hadoop and HDFS data, complex data types, such as arrays, maps, structs, as well as JSON data are more prevalent. Second, the users utilize UDFs (user-defined-functions) very widely to express their business logic, which is sometimes very awkward to express in SQL itself. These two factors complicate the query optimization further in the Hadoop system.

@@

\vspace{-2mm}
\setcounter{mycounter}{0}
\begin{list}{\arabic{mycounter}.}{\usecounter{mycounter}\leftmargin=1.4em}
  \itemsep=0.5em
  \parsep=0.5em
  \parskip=0em
\item {{\bf Native systems with Big Data run-times:} These are systems that utilize a big data run-time such as Tez \cite{tezsigmod15}, or Spark \cite{sparkhotcloud10} to execute SQL queries. This category includes systems such as Hive \cite{hive} and Spark SQL \cite{sparkSQL}.}
\item{ {\bf Native systems with MPP architectures:} These are systems that employ an MPP database architecture that uses stand-alone long-running processing to execute SQL queries over HDFS data. This category includes systems such as Impala \cite{Kornacker+15}, Presto, Drill \cite{drill}, and Big SQL \cite{bigsql}}
\item{{\bf Database hybrids:} These are systems that combine a relational SQL engine with Hadoop run-time. This category includes systems such as HadoopDB \cite{hadoopdb}, HAWQ \cite{hawq}, and Vortex.}
\item{ {\bf Database connectors:} This category includes the extensions provided by major database vendors to reach out to Hadoop and ingest HDFS data.}
\item{ {\bf noSQL/Hbase specific variants:} These are systems that provide a declarative SQL layer on top of noSQL systems, HBase in particular. The systems in this category are Phoenix \cite{phoenix}, and Splice Machine \cite{splice}. Note that these systems focus on OLTP-like workloads, whereas all other categories address OLAP-like analytics}
\end{list}

@@

Database hybrids have emerged to fill this gap, leveraging
pre-existing robust database code for query processing and/or storage
in order to accerlate production-ready implementations of
SQL-on-Hadoop solutions. We will discuss three examples of database
hybrids: HadoopDB~\cite{hadoopdb} which uses large amounts of
PostgreSQL code; Hawq~\cite{hawq} which uses large amounts of
Greenplum code; and Vortex which uses large amounts of Actian
Vectorwise code.

>>>>

It is important to note that there are important distinct characteristics of this emerging eco-system that are different than traditional relational warehouses.
First, in the world of Hadoop and HDFS data, complex data types, such as arrays, maps, structs, as well as JSON data are more prevalent.
Second, the users utilize UDFs (user-defined-functions) very widely to express their business logic, which is sometimes very awkward to express in SQL itself. 
Third, often times there is little control over HDFS. Files can be added or modified outside the tight control of a query engine, making statistics maintenance a challenge.
These factors complicate the query optimization further in the Hadoop system.

@@

\vspace{-2mm}
\setcounter{mycounter}{0}
\begin{list}{\arabic{mycounter}.}{\usecounter{mycounter}\leftmargin=1.4em}
  \itemsep=0.5em
  \parsep=0.5em
  \parskip=0em
\item {{\bf Native systems with Big Data run-times:} These are systems that utilize a big data run-time such as Tez \cite{tezsigmod15}, or Spark \cite{sparkhotcloud10} to execute SQL queries. This category includes systems such as Hive \cite{hive} and Spark SQL \cite{sparkSQL}.}
\item{ {\bf Native systems with MPP architectures:} These are systems that employ an MPP database architecture that uses stand-alone long-running processing to execute SQL queries over HDFS data. This category includes systems such as Impala \cite{Kornacker+15}, Presto \cite{prestofcb}, Drill \cite{drill}, and Big SQL \cite{bigsql}}
\item{{\bf Database hybrids:} These are systems that combine a relational SQL engine with Hadoop run-time. This category includes systems such as HadoopDB \cite{hadoopdb}, HAWQ \cite{hawq}, and Vortex \cite{vortex}.}
\item{ {\bf Database connectors:} This category includes the extensions provided by major database vendors to reach out to Hadoop and ingest HDFS data.}
\item{ {\bf noSQL/Hbase specific variants:} These are systems that provide a declarative SQL layer on top of noSQL systems, HBase in particular. The systems in this category are Phoenix \cite{phoenix}, and Splice Machine \cite{splice}. Note that these systems focus on OLTP-like workloads, whereas all other categories address OLAP-like analytics}
\end{list}

@@

Database hybrids have emerged to fill this gap, leveraging
pre-existing robust database code for query processing and/or storage
in order to accerlate production-ready implementations of
SQL-on-Hadoop solutions. We will discuss three examples of database
hybrids: HadoopDB \cite{hadoopdb} which uses large amounts of
PostgreSQL code; Hawq \cite{hawq} which uses large amounts of
Greenplum code; and Vortex \cite{vortex} which uses large amounts of Actian
Vectorwise code.
